Temporal Evolution of Erosion in Pore-Networks:\\ From Homogenization to Instability

### 2022-04-12 review_response/Response2.tex
BeforePublication

\Question{
The only answer that I do not find completely satisfying is to this
point: ``…since the authors have access to experimental setups, it
would greatly improve this work if they could provide experimental
proof of the phase transition in a real system." To what the authors
answered: `` With regards to the experimental tests, all the authors of
the current manuscript are theorists, and experimental studies of this
work is out of our scope." 2/3 of the authors of this manuscript are coauthors of Reference [31],
which presents experiments related to the results of the current
manuscript. Actually, one of the co-first authors of this manuscript
is also co-first author in [31]. I do not think that all publications
in PRL should include experimental results, but in this particular
case it is striking that the authors do not try to validate their
theory having (apparently) all the experimental tools at hand.
\newline
To summarize, the main and unanimous criticism from the first round of
review was if the paper was physically grounded. On the one hand, I
think the authors have greatly improved the manuscript including more
realistic models of physical erosion. However, on the other hand, I
think the authors have missed the opportunity of including at least
some preliminary experimental results that support their theory or at
least to discuss in their answer why this was not possible.
\newline
Finally, whereas I do not strongly support publication in PRL, I do
not oppose to it either, I leave the final decision to the editors.
\newline}

\Answer{We thank the referee for appreciating the more realistic physical
models utilized in the revised version and the extended theoretical
framework we used.
With regards to the question about the potential experimental tests of
the theory:
We completely agree with the referee that performing experiments to
test our theory is important. Unfortunately performing them is
currently unfeasible for us, since the Amir group is a theory group
and has no lab space.
We would like to make some clarifications with regards to Ref. [31].
In this work Dr. Shima Parsa, then a postdoc in the Weitz lab, had
empirical observations relating the mean velocity and permeability as
flow in a porous material is clogged with a polymer. Dr. Ahmad Zareii,
working under the guidance of Prof. Ariel Amir, set out to explain the
observed scaling using a combination of numerical simulations and
analytical work, which led to a joint publication (Ref. [31] mentioned
by the referee). Note that the experiments were performed by Dr. Parsa
over the course of several years while she was at Harvard. Dr. Parsa
has recently started as faculty at the Rochester Institute of
Technology, and therefore the experimental system associated with Ref.
[31] is no longer active at Harvard. We also resonated with the
comments of Referee D: "Experiments are notoriously hard to perform
and can not be produced easily, especially when the parameter space
that needs to be explored is as wide as in this study (without
considering the 3D and polar simulations in the supplementary). I
believe that the results are convincing, and requiring an experiment
is somewhat unrealistic for the scope of this study. In many ways,
this study can trigger the community to search for this ordering
parameter in existing experimental setups, which will advance the
field further by exploring the role of flow rate to shear rate or mean
pore size." Indeed, we hope that as Referee D mentioned our
theoretical study will trigger further experimental work in the
future.}

@@

\Answer{We thank the reviewer for their thoughtful and thorough comments.}

@@

\Question{Having said that, the work does have a significant weakness in that
the primary model studied is a power law and therefore scale-free. The
scale-free nature is the primary reason why the local dynamics set the
behavior of the entire network. The majority of interesting real-world
cases will have a scale associated with them (e.g. a non-trivial
dependence on flow-rate). In the more general case we will find that
Eq. (3) and the condition described just after it will hold for some
flow rates but not all. This dependence on flow rate (or applied
pressure drop) is not well explained in the manuscript. Also, the new
investigation of a tanh profile shows that intermediate states are
possible - i.e. states that lie somewhere between homogenization and
instability. It would be good for the manuscript to acknowledge some
of these limitations of their result - which remains interesting.\newline}

\Answer{We thank the reviewer for the valuable comment. Indeed our model has limitations. We have now expanded our conclusion to include such limitations as mentioned by the reviewer.

admit scale free after Eq. 3

If not parallel, g(r) can be more complex

acknowledge inbetween cases in the main text ...
}

@@

\Answer{We would like to point out that for $n=3$ the network does not stay the same, since series connections are always moving toward homogenization, while parallel connections are staying in a constant proportion. This results in seeing differences between the statistics of the network at $t=0$ and the evolved network at later stages for $n=3$ (see Fig. 2 in the main text). Additionally, we expect the transition to happen at  approximate $n\approx3$ (see comment 2G for more details). We have now included this point in the main manuscript to avoid confusion.}
%
\\

\Answer{From \p4 of the main text:}
\AnswerQ{
\begin{quote}
    ``It is to be noted that while the flow ratio remains constant for parallel edges, the connection in series are evolving toward homogenization and that is why we find a difference between evolved network for $n=3$ and the network at $t=0$.''
\end{quote}
}

@@

\Answer{We thank the reviewer for their careful comment. We have now included the definition in the main text.}%

\AnswerQ{
\begin{quote}
    ``To understand the transition in network behavior during erosion, we focus on a simplified model with only two tubes \add{of the same length $L$} in parallel or series with a general erosion dynamics as $dr_{i}/dt = f(q_{i},r_{i}), i=1,2$ (Figs. \ref{fig:fig3}b-c)''   
\end{quote}
}

@@

\Answer{We thank the reviewer for their valuable recommendations. The physical origin the scale-free criteria $r_1/r_2$ turns out to be the result of power law dependence of conductance with respect to radius, meaning that our homogenization condition can be generalized to nonlaminar flow as well. The relevant part is attached. }
\AnswerQ{
\begin{quote}
    ``In fact, this condition will hold for nonlaminar flow as long as $C\propto r^k (k > 0)$. The power law dependence of conductance with respect to radius leads to the scale-free criteria $r_1/r_2$. ''   
\end{quote}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%% Review # 3 %%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage 
\vspace{10 mm}
\noindent
\Hline \\
\textbf{Response to Reviewer D} \\
\Hline
\\

\Question{This paper presents 2D and 3D pore-network models of single-phase flow through porous media, involving channel erosion. The erosion rate in a channel is modeled as a function of the flow rate through the channel and the radius of the channel.The manuscript presents an interesting transition in the behaviour of these models from a stabilizing behavior where narrow channels erode more quickly than wider ones (``homogenization'') to an unstable regime where all the flow ends up in a few pathways ``instability''). Overall I judge this to be an interesting paper that may be worth publishing in a journal such as PRFluids, but falls short of the level of impact required for PRL. I explain this below.  The manuscript has many strengths which include: - it's very clearly written and with excellent figures that clearly demonstrate the phenomena being explored; - good numerical modeling that includes 2D and 3D simulations in random and regular network topologies; - results which are universal within the class of models being considered here, being valid 2D and 3D and in regular and random networks; - a clear theory to explain the results using a local analysis,suggesting links between global and local dynamics even in quite complex systems; - links to flow in biological systems
\newline}

\Answer{We thank the reviewer for their careful reading of the manuscript and their constructive
remarks.}

\vspace{0.5cm}
\textbf{Comment 3A}
\noindent \vspace{-0.2cm}\\ \Hline\\
\Question{The weaknesses in the work include: the physical relevance of the model is not clear, except for the specific biological case ... more specifics in studies of models that are closest to this one are needed, particularly also mentioning the outcomes of those studies. Some of the model's (apparent) weaknesses are being a pure power law, its scaling with pressure is unphysical: the erosion rate in all channels merely increases uniformly with applied pressure, meaning that erosion occurs even at near-zero speed/stress; there is no onset threshold.\newline}

\Answer{We appreciate the reviewers comment. The concern here is similar to comment 2F of Reviewer 2, and we repeat our reply here. Erosion in saturated granular materials occurs when the shear stress at the wall exceeds the cohesive strength of the solid matter. In the revised version, we study more realistic models including: (i) a model where the onset of erosion happens after reaching a threshold value; (ii) a model with a different (nonlinear) functional dependence on shear stress, where erosion starts after a certain threshold and eventually saturates to a maximum value which is determined by the maximal detachment rate of particles at the tube's boundary. We make predictions for both models using our analysis and find that the analytical results agree well with the numerical simulations. These results illustrate the generality of our model in relating to various physical models. (iii) Additionally, we expand our discussion before introducing the model to make a better connection with relevant studies of erosion. 
We believe that these new results illustrate that the phase transition is a general effect, and that our analysis using the local dynamics model has the predictive power of identifying the phase transition condition. In the following we attach the relevant excerpts of the main text and the supplementary material.}

>>>>

\Question{I have read the three referee reports
and the authors' answers to all questions and criticisms. I think the
authors have tried to answer all the referee's points and most answers
are satisfactory to me.}

\vspace{0.3 cm}

\Answer{We appreciate the referee's thorough reviews.}
\vspace{0.3 cm}

\Question{
The only answer that I do not find completely satisfying is to this
point: ``...since the authors have access to experimental setups, it
would greatly improve this work if they could provide experimental
proof of the phase transition in a real system." To what the authors
answered: `` With regards to the experimental tests, all the authors of
the current manuscript are theorists, and experimental studies of this
work is out of our scope." 2/3 of the authors of this manuscript are coauthors of Reference [31],
which presents experiments related to the results of the current
manuscript. Actually, one of the co-first authors of this manuscript
is also co-first author in [31]. I do not think that all publications
in PRL should include experimental results, but in this particular
case it is striking that the authors do not try to validate their
theory having (apparently) all the experimental tools at hand.
\newline
To summarize, the main and unanimous criticism from the first round of
review was if the paper was physically grounded. On the one hand, I
think the authors have greatly improved the manuscript including more
realistic models of physical erosion. However, on the other hand, I
think the authors have missed the opportunity of including at least
some preliminary experimental results that support their theory or at
least to discuss in their answer why this was not possible.
\newline
Finally, whereas I do not strongly support publication in PRL, I do
not oppose to it either, I leave the final decision to the editors.
\newline}

\Answer{We thank the referee for appreciating the more realistic physical
models utilized in the revised version and the extended theoretical
framework we used.
With regards to the question about the potential experimental tests of
the theory:
We completely agree with the referee that performing experiments to
test our theory is important. Unfortunately performing them is
currently unfeasible for us, since the Amir group is a theory group
and has no lab space.
We would like to make some clarifications with regards to Ref. [31].
In this work Dr. Shima Parsa, then a postdoc in the Weitz lab, had
empirical observations relating the mean velocity and permeability as
flow in a porous material is clogged with a polymer. Dr. Ahmad Zareei,
working under the guidance of Prof. Ariel Amir, set out to explain the
observed scaling using a combination of numerical simulations and
analytical work, which led to a joint publication (Ref. [31] mentioned
by the referee). Note that the experiments were performed by Dr. Parsa
over the course of several years while she was at Harvard. Dr. Parsa
has recently started as faculty at the Rochester Institute of
Technology, and therefore the experimental system associated with Ref.
[31] is no longer active at Harvard. We hope that as Referee D mentioned our theoretical study will trigger further experimental work in the future.}

@@

\Answer{We thank the reviewer for their thoughtful comments.}

@@

\Question{Having said that, the work does have a significant weakness in that
the primary model studied is a power law and therefore scale-free. The
scale-free nature is the primary reason why the local dynamics set the
behavior of the entire network. The majority of interesting real-world
cases will have a scale associated with them (e.g. a non-trivial
dependence on flow-rate). In the more general case we will find that
Eq. (3) and the condition described just after it will hold for some
flow rates but not all. This dependence on flow rate (or applied
pressure drop) is not well explained in the manuscript. Also, the new investigation of a tanh profile shows that intermediate states are
possible - i.e. states that lie somewhere between homogenization and
instability. It would be good for the manuscript to acknowledge some
of these limitations of their result - which remains interesting.
\newline}

\Answer{We thank the reviewer for the valuable comment. Indeed for a more general erosion dynamics the condition in Eq.~(3) might not hold for all the local connections and as noted a transitional region is observed; in other words, the stability criterion is obeyed initially for some pipes (below a threshold shear stress/radius) but does not hold above it. Naively, one would expect flow in the stable pipes to be maintained over time, while channelization would occur for the others. However, during the process of channelization, flow is redistributed throughout the network and the flow distribution therefore changes -- implying that some pipes that were initially in the stable regime may become unstable at later times. This is indeed what we observe in our simulations, for non-scale-free erosion laws. Furthermore, it is precisely this dynamics which enables the formation of \emph{multiple} channels, in contrast to the single channels observed for the scale-free erosion laws.

We have now included the remarks from the supplementary material in the main text. With regards to addressing the limitations, we have now expanded our conclusion to address the limitations of our model. The relevant parts from the main text are included here for completeness.\newline}
% our model has limitations
% . We have now expanded our conclusion to include such limitations as mentioned by the reviewer. admit scale free after Eq. 3 If not parallel, g(r) can be more complex acknowledge inbetween cases in the main text ...

\Answer{From \p3 of the main text:}
\AnswerQ{
\begin{quote}
   ``Note that, the fact that local dynamics predict the fate of the overall network in the way described above is a result of the scale-free nature of the aforementioned power-law erosion model. Later, we will discuss richer phenomenology that may arise in the case of physical erosion laws possessing an intrinsic scale.''
\end{quote}
}
\Answer{From \p4 of the main text:}
\AnswerQ{
\begin{quote}
    ``It is to be noted that the scale-free nature of our model as discussed after Eq. (3) is the main reason for the local dynamics setting the behavior of the whole network. Physical systems often have a scale associated with them, and in such cases, Eq.~(3) and the condition described just after hold true for some flow rates but not all. Investigating nonlinear erosion dynamics, we found that transitional states are then possible.''
\end{quote}
\AnswerQ{
\begin{quote}
    ``Note that for general erosion dynamics, the condition in Eq. (3) might hold for some but not all of the pipes at a given time. Furthermore, during the process of erosion, flow is redistributed throughout the network and therefore the flow distribution changes. This dynamical process enables the formation of \emph{multiple} channels, in contrast to the single channels observed for the scale-free erosion laws (see SM \S7 and Fig. S10).''
\end{quote}
}
}

@@

\Answer{We would like to thank the reviewer for their thorough reading. In fact, moving from $\dot{m}$ to $\dot{r}$, a factor of $1/r$ is introduced, since $dm=2\pi \rho rl dr$ in a cylindrical pipe. There is a typo in this equation and the factor should read $\kappa/r$. We have corrected the equation and additionally included more information to make the derivation clearer. The reference for the shear stress proportionality with $q/r^3$ is now added to the main text. The relevant part of the manuscript is included below.}
%
\\

\Answer{From \p2 of the main text:}
\AnswerQ{
\begin{quote}
    ``In a class of erosion models previously studied [34-36] the eroded mass per unit area is linearly proportional to the excess shear, i.e.,  $\dot{m}  = \kappa  (\tau_w-\tau_c)$ where $\kappa$ is the linear proportionality constant. This erosion model effectively results in an increase in the radius of the edges in our network model where $\dot{r} \propto \kappa/r (|q|/r^3 - \tau_c)$ , where the $1/r$ factor comes from the eroded mass $dm =2\pi\rho r l dr$ for a cylindrical pipe and $|q|/r^3$ comes from the shear stress of laminar flow in pipes [37].''
\end{quote}
}

@@

\Answer{We thank the reviewer for the comment. We have now included the definition in the main text.}%

\Answer{From \p3 of the main text:}
\AnswerQ{
\begin{quote}
    ``To understand the transition in network behavior during erosion, we focus on a simplified model with only two tubes of the same length $L$ in parallel or series with a general erosion dynamics as $dr_{i}/dt = f(q_{i},r_{i}), i=1,2$.''   
\end{quote}
}

\vspace{0.5cm}
\textbf{Comment C5}
\noindent \vspace{-0.2cm}\\ \Hline\\

@@

\Answer{We thank the reviewer for their valuable suggestion. The physical origin of the scale-free criteria $r_1/r_2$ turns out to be a result of power-law dependence of conductance with respect to radius (i.e., $C\propto r^k$). Interestingly, this indicates that our homogenization condition can be generalized to non-laminar flow where conductance has again a power-law dependence on the radius [50]. The relevant part is attached here for completeness. \newline }
\\

\Answer{From \p3 of the main text:}
\AnswerQ{
\begin{quote}
    ``Note that the scale-free criteria here (i.e., the dependence on the dimensionless factor $r_1/r_2$) results from the power-law dependence of conductance on radius, i.e., $C\propto r^k (k > 0)$. As a result, in addition to laminar flow (where $C\propto r^4$), the above condition also holds true for nonlaminar flow examples such as molecular flow where $C\propto r^3$ [50].''
    % ``In fact, this condition will hold for nonlaminar flow as long as . The power law dependence of conductance with respect to radius leads to the scale-free criteria $r_1/r_2$. ''
\end{quote}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%% Review # 3 %%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-04-12 supp.tex
BeforePublication

This condition guarantees that at each step a small amount of material is eroded and there is no sudden change in the network. We further test the convergence by decreasing $\max(\Delta r_{ij})$ to half and we observe that the average relative change in the flux vector is $\approx 1.2\%$, and the PDFs remain intact without any notable change. The network size for the 2d/3d networks is $N_x=100,N_y=50$/$N_x=50,N_y=12, N_z=12$ unless mentioned otherwise. In either case, the network has an order of $10^3$ edges. \add{Note that the physical assumption of low Reynolds regime Re$\ll$1 at all times ensures that the Poisoulle flow assumption remains valid. Additional dimensionless transport numbers such as the P\'eclet number (ratio of advection to diffusion) or the Damkohler number (ratio of chemical reaction rate to the transport phenomena rate) are assumed to be respectively very large and very small, resulting in advection to be much stronger than diffusion and chemical reaction.  It is to be noted that considering diffusion or chemical reactions might change the dynamics of erosion as it requires tracking a secondary concentration field that might affect the local erosion law.} The code is publicly available in a GitHub repository \cite{githubrepo}.

@@

In order to check the robustness of our results with respect to the topology of the network, we run our simulations on a three-dimensional random network with Voronoi cell initialization of nodes in space (Fig. \ref{SIfig:fig2-3d}), and also for a two dimensional topologically ordered network with a diamond grid (Fig. \ref{fig:fig2-diamond}), where  both networks are initialized with a uniform random distribution for the diameter of the tubes. We find that regardless of the topology of the network, an erosion dynamics with $n>3$ results in a homogenized network, while $n<3$ results in the channelization instability. We further check the effect of initial randomness on the fate of the network. We use a two-dimensional diamond-grid network with a narrow uniform initial distribution of tube diameters around an average diameter $d_0$ with only a very small variation ($3\%$), i.e., the tube diameters are sampled from a uniform distribution with $\mathcal{U}(d_0(1-\epsilon), d_0(1+\epsilon))$. We again find that for the networks with $n<3$ channels are formed, while for $n>3$ the network stays homogenized. \add{In addition to the network topology and initial condition, we test the robustness of our results to changes in the boundary condition. We choose a circular domain, with the central node being the inlet and all the nodes at the outer boundary being outlets. Similar to the transition reported for the square

@@

     \caption{\textcolor{blue}{Erosion in a topologically random 3D network of tubes with a circular topology with $N = 3000 $ randomly distributed nodes and an initial uniform broad distribution of tube diameters randomly sampled from a uniform distribution with $\mathcal{U}(1,14)$. Shown are snapshots of the network, PDF of normalized fluid flux $q/\langle q \rangle$, and normalized edge radius distribution $r/\langle r \rangle$ at the initial time $t=0$, and also after $N$ erosion steps for different powers of erosion $n$. We stop the erosion after $N$ steps such that $\langle r\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion.}}\label{SIfig:fig2-3d_circular}

@@

\clearpage
\section{Erosion with a threshold}
%
\add{In order to make our erosion model more realistic, we modify the model as  
%
\begin{align}
    \frac{dr}{dt } = \begin{cases} q^m/r^n - \beta & q^m/r^n\geq \beta\\
    0 & q^m/r^n<\beta
    \end{cases} \label{eq:thresh}
\end{align}
%
where $\beta$ is the threshold for the onset of erosion. We pick $\beta = 0.25\langle q^m/r^n\rangle $ and $\beta = 0.9 \langle q^m/r^n\rangle$ corresponding to a small and a large threshold value. In both cases, we ran our simulations for different values of $m$ and $n$, and found that different behaviors of homogenization and channelization persists (see Fig. \ref{fig:figthresh1}). The result can be explained by applying the general condition for the phase transition as discussed in the main text. Following the same argument, it can be shown that the homogenization condition (Eq. 3 in the main text) is equivalent to the condition 
%
\begin{align}
4m-n-1 < -\frac{\beta}{q^m/r^n}.    
\end{align}
%
which can be found using $g(r) = ({\pi \Delta p}/{8\mu L})^{m} r^{m-n} - \beta $ assuming a constant pressure boundary condition. Note that the condition above depends on the value of $q^m/r^n$ which varies over the network edges. Taking the mean value of $q^m/r^n$ to find an average value for the transition condition, the transition line becomes $4m-n=0.75$ and $4m-n=0.1$ for the small and large thresholding value respectively. As a result the thresholding condition slightly shifts the intercept of the transition line while the slope remains the same. In Fig.~\ref{fig:figthresh1}, we plot the new transition lines, where the prediction for the phase boundaries matches well with the simulation results.}

\begin{figure}[htp]
    \centering
 \includegraphics[width=0.45\textwidth]{./Fig_thres0.25.pdf}
  \includegraphics[width=0.45\textwidth]{./Fig_thresh0.9.pdf}
     \caption{\add{Evolution of a randomly initialized network for various powers of $m$ and $n$ in Eq. \eqref{eq:thresh} with a small threshold value of $\beta = 0.25\langle q^m/r^n\rangle $ (left figure), and a large threshold value of $\beta = 0.9\langle q^m/r^n\rangle$ at $t=0$. The network is randomly initialized with $50\times 50$ randomly distributed pores. The black line shows the theoretical transition boundary between channelization instability and homogenization obtained using local dynamics model, i.e., $4m-n=0.75$ (left figure), and $4m-n=0.1$ (right figure).} }\label{fig:figthresh1}% 
\end{figure}

@@

\add{\section{General shear dependent erosion model}}
\add{

@@

     \caption{\textcolor{blue}{(a) A 2D random network with $N_x=50$ and $N_y=50$ points. (b) The probability distribution function (PDF) of the shear stress for the edges of the network. (c)/(f)/(i) The PDF of the shear stress along with nonlinear erosion law $f(\tau)$ (solid red line) and $\tau f'(\tau)$ (dashed red line). The areas where the homogenization condition (i.e., $f(\tau)>\tau f'(\tau)$) is satisfied are shaded with green color, otherwise, the channelization condition is satisfied and the region is shaded with red color. (d)/(g)/(j) Snapshots of the 2D network after erosion using the nonlinear erosion law shown in the previous row until the average radius of the network is increased to twice its original size. (e)/(h)/(k) The PDF of the shear stress at the walls for the final snapshot of the network shown in the previous row. 
    }}

@@

\begin{thebibliography}{8}%

@@

@@

>>>>

This condition guarantees that at each step a small amount of material is eroded and there is no sudden change in the network. We further test the convergence by decreasing $\max(\Delta r_{ij})$ to half and we observe that the average relative change in the flux vector is $\approx 1.2\%$, and the PDFs remain intact without any notable change. The network size for the 2d/3d networks is $N_x=100,N_y=50$/$N_x=50,N_y=12, N_z=12$ unless mentioned otherwise. In either case, the network has an order of $10^3$ edges. {Note that the physical assumption of low Reynolds regime Re$\ll$1 at all times ensures that the Poisoulle flow assumption remains valid. Additional dimensionless transport numbers such as the P\'eclet number (ratio of advection to diffusion) or the Damkohler number (ratio of chemical reaction rate to the transport phenomena rate) are assumed to be respectively very large and very small, resulting in advection to be much stronger than diffusion and chemical reaction.  It is to be noted that considering diffusion or chemical reactions might change the dynamics of erosion as it requires tracking a secondary concentration field that might affect the local erosion law.} The code is publicly available in a GitHub repository \cite{githubrepo}.

@@

In order to check the robustness of our results with respect to the topology of the network, we run our simulations on a three-dimensional random network with Voronoi cell initialization of nodes in space (Fig. \ref{SIfig:fig2-3d}), and also for a two dimensional topologically ordered network with a diamond grid (Fig. \ref{fig:fig2-diamond}), where  both networks are initialized with a uniform random distribution for the diameter of the tubes. We find that regardless of the topology of the network, an erosion dynamics with $n>3$ results in a homogenized network, while $n<3$ results in the channelization instability. It is to be noted that the different distributions obtained for different powers of $n$ represent different statistics that cannot be collapsed. 
We further check the effect of initial randomness on the fate of the network. We use a two-dimensional diamond-grid network with a narrow uniform initial distribution of tube diameters around an average diameter $d_0$ with only a very small variation ($3\%$), i.e., the tube diameters are sampled from a uniform distribution with $\mathcal{U}(d_0(1-\epsilon), d_0(1+\epsilon))$. We again find that for the networks with $n<3$ channels are formed, while for $n>3$ the network stays homogenized. {In addition to the network topology and initial condition, we test the robustness of our results to changes in the boundary condition. We choose a circular domain, with the central node being the inlet and all the nodes at the outer boundary being outlets. Similar to the transition reported for the square

@@

     \caption{Erosion in a topologically random 3D network of tubes with a circular topology with $N = 3000 $ randomly distributed nodes and an initial uniform broad distribution of tube diameters randomly sampled from a uniform distribution with $\mathcal{U}(1,14)$. Shown are snapshots of the network, PDF of normalized fluid flux $q/\langle q \rangle$, and normalized edge radius distribution $r/\langle r \rangle$ at the initial time $t=0$, and also after $N$ erosion steps for different powers of erosion $n$. We stop the erosion after $N$ steps such that $\langle r\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion.}\label{SIfig:fig2-3d_circular}

@@

\clearpage
\section{Order parameter and Phase transition}

To quantify the transition of the network between the channeling instability and homogenization, we define an order parameter~\footnote{The order parameter defined here is inspired by the participation ratio (PR) employed to quantify the localization of an eigenstate in the analysis of Anderson localization~\cite{kramer1993localization}.} % $\mathcal{O}$ 
		%
		\begin{align}
			\mathcal{O} = \frac{1}{N-1} \left( N - \frac{\left( \sum_{ij} q^2_{ij}\right)^2}{ \sum_{ij} q^4_{ij}}\right),
		\end{align}
		%
where $N$ is the number of edges.  The order parameter $\mathcal{O}=0 $ when the flux through every edge is identical. On the other hand, when fluid flux becomes highly localized with only a few edges with non-zero flux, $\mathcal{O}\to 1$.  We numerically calculated the  order parameter $\mathcal{O}$ for {randomly initialized networks, averaged over 20 different realization} and the results are shown in Fig. \ref{fig:fig3-SIa}.  We find that at $n\approx 3$ the order parameter remains unchanged; however for $n>3$ the order parameter moves toward zero, where the flow becomes more uniform, and for $n<3$ the order parameter goes toward unity, where channels are developed. This indicates a phase transition at $n \approx 3$ in the long-time behavior of the network.

% To quantify the transition of the network between the channeling instability and homogenization, we numerically calculated the order parameter $\mathcal{O}$ for randomly initialized networks, averaged over 20 different realization and the results are shown in Fig. \ref{fig:fig3-SI}a.  The results indicate a phase transition at $n\approx 3$ where at $n\approx 3$ the order parameter remains unchanged; while for $n > 3$ the order parameter moves toward zero and the flow becomes more uniform, and for $n < 3$ the order\ parameter goes toward unity and channels are developed.  Additionally, as discussed in section \textit{Local Dynamics Model} of the main text, to understand the transition in network behavior during erosion, we study a simplified model with only two tubes, where the results are illustrated in Fig.~\ref{fig:fig3-SI}b-c. Assuming two cylindrical tubes in series (Fig. \ref{fig:fig3-SI}b), the normalized pressure at the middle junction between the tubes approaches the average of pressure at both sides of the tube resulting in a homogenized pressure distribution. Contrary to tubes in series, when the tubes are in parallel, depending on the erosion dynamics, different behaviors of homogenization, channelization, or staying the same can be observed.

\begin{figure}[!h]
	% \centering
	\includegraphics[width = 0.55\textwidth]{Fig3-order.pdf}
	\caption{ Order parameter $\mathcal{O}$ calculated from simulation results presented in Fig. 2 of the main text for different powers of $n$ with $m=1$ plotted over time.  
    % 	Local Dynamics Model: Tubes in series (a) or parallel (b) configuration. The tube radius dynamically change with the erosion law, i.e., $f(q,r) = \alpha |q|^m/r^n$. When the tubes are in series, for any $m,n$, (a$_1$) the normalized pressure at the middle junction between the tubes $\tilde p_m = (p_m - p_l)/(p_r-p_l)$ approaches $1/2$ which results in a homogenized pressure distribution. When the tubes are in parallel, for (b$_1$) $m=(n+1)/4$, the flow ratio between the pipes does not change over time; (b$_2$)  $m<(n+1)/4$ the flow distributes between the tubes equally and  $q_1/(q_1+q_2) \to 1$ which results in the homogenization of the network; (b$_3$) $m>(n+1)/4 $, the entire flow eventually passes through one of the tubes, and channeling occurs. 
	}\label{fig:fig3-SIa}
\end{figure}

\section{Local Dynamics Model}
As discussed in section \textit{Local Dynamics Model} of the main text, to understand the transition in network behavior during erosion, we study a simplified model with only two tubes, where the results are illustrated in Fig.~\ref{fig:fig3-SIb}a-b. Assuming two cylindrical tubes in series (Fig. \ref{fig:fig3-SIb}a), the normalized pressure at the middle junction between the tubes approaches the average of pressure at both sides of the tube resulting in a homogenized pressure distribution. Contrary to tubes in series, when the tubes are in parallel (Fig.~\ref{fig:fig3-SIb}b), depending on the erosion dynamics, different behaviors of homogenization, channelization, or staying the same can be observed.

\begin{figure}[!h]
	% \centering
	\includegraphics[width = 0.95\textwidth]{Fig3-model.pdf}
	\caption{Local Dynamics Model: Tubes in series (a) or parallel (b) configuration. The tube radius dynamically change with the erosion law, i.e., $f(q,r) = \alpha |q|^m/r^n$. When the tubes are in series, for any $m,n$, (a$_1$) the normalized pressure at the middle junction between the tubes $\tilde p_m = (p_m - p_l)/(p_r-p_l)$ approaches $1/2$ which results in a homogenized pressure distribution. When the tubes are in parallel, for (b$_1$) $m=(n+1)/4$, the flow ratio between the pipes does not change over time; (b$_2$)  $m<(n+1)/4$ the flow distributes between the tubes equally and  $q_1/(q_1+q_2) \to 1$ which results in the homogenization of the network; (b$_3$) $m>(n+1)/4 $, the entire flow eventually passes through one of the tubes, and channeling occurs. 
	}\label{fig:fig3-SIb}
\end{figure}

\section{Erosion with a threshold}
%
{In order to make our erosion model more realistic, we modify the model as  
%
\begin{align}
    \frac{dr}{dt } = \begin{cases} q^m/r^n - \beta & q^m/r^n\geq \beta\\
    0 & q^m/r^n<\beta
    \end{cases} \label{eq:thresh}
\end{align}
%
where $\beta$ is the threshold for the onset of erosion. We pick $\beta = 0.25\langle q^m/r^n\rangle $ and $\beta = 0.9 \langle q^m/r^n\rangle$ corresponding to a small and a large threshold value. In both cases, we ran our simulations for different values of $m$ and $n$, and found that different behaviors of homogenization and channelization persists (see Fig. \ref{fig:figthresh1}). The result can be explained by applying the general condition for the phase transition as discussed in the main text. Following the same argument, it can be shown that the homogenization condition (Eq. 3 in the main text) is equivalent to the condition 
%
\begin{align}
4m-n-1 < -\frac{\beta}{q^m/r^n}.    
\end{align}
%
which can be found using $g(r) = ({\pi \Delta p}/{8\mu L})^{m} r^{m-n} - \beta $ assuming a constant pressure boundary condition. Note that the condition above depends on the value of $q^m/r^n$ which varies over the network edges. Taking the mean value of $q^m/r^n$ to find an average value for the transition condition, the transition line becomes $4m-n=0.75$ and $4m-n=0.1$ for the small and large thresholding value respectively. As a result the thresholding condition slightly shifts the intercept of the transition line while the slope remains the same. In Fig.~\ref{fig:figthresh1}, we plot the new transition lines, where the prediction for the phase boundaries matches well with the simulation results.}

\begin{figure}[htp]
    \centering
 \includegraphics[width=0.45\textwidth]{./Fig_thres0.25.pdf}
  \includegraphics[width=0.45\textwidth]{./Fig_thresh0.9.pdf}
     \caption{Evolution of a randomly initialized network for various powers of $m$ and $n$ in Eq. \eqref{eq:thresh} with a small threshold value of $\beta = 0.25\langle q^m/r^n\rangle $ (left figure), and a large threshold value of $\beta = 0.9\langle q^m/r^n\rangle$ at $t=0$. The network is randomly initialized with $50\times 50$ randomly distributed pores. The black line shows the theoretical transition boundary between channelization instability and homogenization obtained using local dynamics model, i.e., $4m-n=0.75$ (left figure), and $4m-n=0.1$ (right figure). }\label{fig:figthresh1}% 
\end{figure}

@@

\section{General shear dependent erosion model}
{

@@

     \caption{(a) A 2D random network with $N_x=50$ and $N_y=50$ points. (b) The probability distribution function (PDF) of the shear stress for the edges of the network. (c)/(f)/(i) The PDF of the shear stress along with nonlinear erosion law $f(\tau)$ (solid red line) and $\tau f'(\tau)$ (dashed red line). The areas where the homogenization condition (i.e., $f(\tau)>\tau f'(\tau)$) is satisfied are shaded with green color, otherwise, the channelization condition is satisfied and the region is shaded with red color. (d)/(g)/(j) Snapshots of the 2D network after erosion using the nonlinear erosion law shown in the previous row until the average radius of the network is increased to twice its original size. (e)/(h)/(k) The PDF of the shear stress at the walls for the final snapshot of the network shown in the previous row. 

@@

\begin{thebibliography}{10}%

@@

\bibitem [{Note1()}]{Note1}%
  \bibinfo {note} {The order parameter defined here is inspired by the
  participation ratio (PR) employed to quantify the localization of an
  eigenstate in the analysis of Anderson localization~\cite
  {kramer1993localization}.}\BibitemShut {Stop}%

@@

\bibitem [{\citenamefont {Kramer}\ and\ \citenamefont
  {MacKinnon}(1993)}]{kramer1993localization}%
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {B.}~\bibnamefont
  {Kramer}}\ and\ \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {MacKinnon}},\ }\bibfield  {title} {\bibinfo {title} {Localization: theory
  and experiment},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {Reports on Progress in Physics}\ }\textbf {\bibinfo {volume} {56}},\
  \bibinfo {pages} {1469} (\bibinfo {year} {1993})}\BibitemShut {NoStop}%
